---
title: "PUT YOUR TITLE HERE"
author: "PUT YOUR NAME HERE"
output:
  html_document:
    df_print: paged
---

**BELOW ARE INSTRUCTIONS FOR HOW TO COMPLETE THE ASSIGNMENT;
MAKE SURE THAT YOU DELETE ALL OF THE COMMENTS IN ALL CAPS
(AS WELL AS THIS COMMENT), BEFORE SUBMITTING**

THIS FIRST CHUNK SHOULD LOAD ALL OF THE PACKAGES (I THINK
YOU WILL NOT NEED TO CHANGE THIS)

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(ggmap)
library(viridis)
library(forcats)
```

THE SECOND CHUNK SHOULD LOAD THE DATASET AND PRODUCE ANY
FILTERED DATASETS THAT YOU NEED IN THE REST OF THE REPORT

```{r}
tracts <- read_csv("https://raw.githubusercontent.com/statsmaths/stat_data/gh-pages/acs_tract_data_2016.csv.gz")
```


```{r}
tracts <- read_csv("https://raw.githubusercontent.com/statsmaths/stat_data/gh-pages/acs_tract_data_2016.csv.gz")
  tracts_lakecounty <- filter(tracts, state=="Indiana", county == "Lake County")
```


```{r}
tracts <- read_csv("https://raw.githubusercontent.com/statsmaths/stat_data/gh-pages/acs_tract_data_2016.csv.gz")
  tracts_lakecounty$ctime_60_up <- tracts_lakecounty$ctime_60_89 + tracts_lakecounty$ctime_90_99 
  
```



```{r}
ggplot(data=tracts_lakecounty, aes(age_00_17, commute_car))+
  geom_point(alpha = .5)+
  geom_point(aes(color = race_white))+
  annotate("rect", xmin=17, xmax=28, ymin=89, ymax=99, 
           fill = "red", alpha= .25)+
  annotate("text",x=19.8, y=100,
           label = "Majority of Commuting White Youths")+
  xlab("Majority Ages 17 and Under")+
  ylab("Majority Commute by Car")+
  ggtitle("Commuting Preferences by Age and Race")
  
```



```{r}
tracts <- read_csv("https://raw.githubusercontent.com/statsmaths/stat_data/gh-pages/acs_tract_data_2016.csv.gz")
  tracts_lakecounty$income_100_up <- tracts_lakecounty$income_100_125 + tracts_lakecounty$income_125_150 + tracts_lakecounty$income_150_200 + tracts_lakecounty$income_200_999
```


```{r}
tracts <- read_csv("https://raw.githubusercontent.com/statsmaths/stat_data/gh-pages/acs_tract_data_2016.csv.gz")
  tracts_lakecounty$income_50_below <- tracts_lakecounty$income_045_050 + tracts_lakecounty$income_040_045 + tracts_lakecounty$income_035_040 + tracts_lakecounty$income_030_035 + tracts_lakecounty$income_025_030 + tracts_lakecounty$income_020_025 + tracts_lakecounty$income_015_020 + tracts_lakecounty$income_010_015 + tracts_lakecounty$income_000_010

```



```{r}
ggplot(data=tracts_lakecounty, aes(population, income_50_below))+
  geom_point()+
  geom_point(aes(color = cut(race_black, 5)))+
  annotate("rect", xmin=700, xmax=3200, ymin=60, ymax=90, 
           fill = "red", alpha= .25)+
  annotate("text",x=6750, y=83,
           label = "Smaller, Majority Black Communities Earning Below $50k", color = "red")+
  xlab("Population")+
  ylab("Income 50k or Under")+
  ggtitle("Racial Distinctions in Income Level")
```



```{r}
tracts <- read_csv("https://raw.githubusercontent.com/statsmaths/stat_data/gh-pages/acs_tract_data_2016.csv.gz")
  tracts_lakecounty$rent_by_income_30_up <- tracts_lakecounty$rent_by_income_30_35 + tracts_lakecounty$rent_by_income_35_40 + tracts_lakecounty$rent_by_income_40_50 + tracts_lakecounty$rent_by_income_50_99
```



```{r}
ggplot(tracts_lakecounty, aes(race_black, rent_by_income_30_up)) +
  geom_point(aes(color = bin(rent_by_income_30_up, 3)))+
  xlab("Percent of Population Identifying as Black")+
  ylab("Percentage of Income Spent on Rent")+
  ggtitle("Top Tier Spenders of Income on Rent Independent of Race")

```




```{r}
rm(tracts_lake, tracts_lake_commute1)
```




NOW, YOUR REPORT SHOULD GO HERE. FEEL FREE TO SPLIT IT UP
WITH SECTIONS `##`, BUT THIS IS NOT REQUIRED. ADD CODE CHUNKS
TO PRODUCE GRAPHICS WHERE THEY BELONG IN THE OUTPUT

one interesting dynamic may be the proximity of this town to chicago, but still outside illinois borders

## References

IF YOU CITE ANY EXTERNAL SOURCES INCLUDE PARANTHETICAL
REFERENCES [SUCH AS: (Arnold, 2017)] IN THE TEXT AND
ADD CITATIONS HERE.


