---
title: "A Statistical Analysis of Gender Inequality in the Film Industry"
author: "Chris Brodsky"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(ggmap)
library(viridis)
library(forcats)
library(smodels)
```


```{r}
actor <- read_csv("https://raw.githubusercontent.com/URstats/stat209-s18-cbrodsky/master/actor.csv")
actor
```


## Introduction

  This project will examine a dataset of actors and the number of projects and academy award nominations for each. The following analysis includes two mean inferences examining each actor's film history and oscar nominations in relation to their gender. Further, the analysis includes a linear regression examining the relationship between the two numeric variables, projects and nominations.  
  My hypothesis is that a statistically significant relationship exists between an actor's gender and the number of projects he or she has worked on, as well as the number of nominations they have received. I also assume there is a clear relationship between those actors who have worked on a larger number of film projects, and the number of Oscar nominations they have received. 
  
## Methods

  For this project, I gathered the relevant data from the IMDB website which includes a filmography section for each actor as well as information pertaining to his or her nomination record. The number of nominations includes those which resulted in a win for each actor. The actors included in the dataset were selected after examining lists of the oscar nominees over the past several years. 
  

## Results

  The first inference and graphic below show the number of film projects actors have worked on as a function of each actors' gender. That the confidence interval values for "gendermale" are both positive indicates a statistically significant difference between the mean number of film projects that male and female actors have worked on.
  The boxplot below illustrates this difference graphically. The median number of male actor projects is approximately 75 compared to the female median of approximately 60. Further, the male category includes several outliers above the fourth quartile ranging from just above 150 to over 200 projects. 

```{r}
model <- lm_basic(number_acting_projects ~ 1 + gender, data = actor)
reg_table(model, level = .9)
```


```{r}
ggplot(actor, aes(gender, number_acting_projects))+
  geom_boxplot()+
  ggtitle("Boxplot: Male and Female Film Projects")+
  xlab("Gender")+
  ylab("Number of Film Projects")
```

  The second output and graphic below examine the difference in the number of academy award (oscar) nominations between male and female actors. Here, the mean inference shows no statistically significant difference in the mean values of nominations among male and female actors. 
  As the boxplot graphic illustrates, the two medians hover closely together just below the 2.5 mark. Likewise, the fourth quartile ranges extend to approximately 7.5 for both categories. 

```{r}
model <- lm_basic(academy_award_noms ~ 1 + gender, data = actor)
reg_table(model, level = .9)
```


```{r}
ggplot(actor, aes(gender, academy_award_noms))+
  geom_boxplot()+
  ggtitle("Boxplot: Male and Female Academy Award Nominations")+
  xlab("Gender")+
  ylab("Number of Oscar Nominations")
```

  The final output and graphic below examine the relationship between the two numeric variables: number of acting projects and academy award nominations. The following graphic illustrates a line of best fit running through the spread of data points for each actor. Although slight, there is a clear linear relationship between the two variables.  


```{r}
model <- lm_basic(number_acting_projects ~1 + academy_award_noms, data = actor)
  reg_table(model, level = .9)
```

```{r}
qplot(number_acting_projects, academy_award_noms, data = actor) + 
  geom_smooth(method = "lm")+
  ggtitle("Relationship Between Number of Projects and Award Nominations")+
  xlab("Number of Film Projects")+
  ylab("Number of Oscar Nominations")
```


## Conclusions

  The results of the inferences above prove my first hypothesis accurate: that a statistically significant difference exists between the mean number of projects worked on by male and female actors. The data shows that men have, on average, worked on a greater number of film projects than their female counterparts. My second hypothesis, however, was proven incorrect: there was in fact no significant difference in the number of oscar nominations between male and female actors.
  This project elaborates on some of the dynamics underlying the recent "Me Too" movement which has, in part, scrutinized inequality in today's film industry. Although the data disproved my second hypothesis, the existence of a statistical difference in the number of film projects between male and female actors demonstrates a potential disparity between the availability of opportunites for women in the industry. Although many women received oscar nominations regardless of the number of projects on their resume, the linear regression model above shows that there is in fact a clear relationship between the two variables. My third hypothesis proved accurate which suggests as opportunities for women in film increase, so too, according to the data, will female oscar nominations increase. 

## References
http://www.imdb.com/ 



